IRS County-to-County Migration Data

### 2018-08-13 manuscript_file.Rmd
minor edits.

To encourage and facilitate the use of this tremendous migration resource, I flattened the IRS migration data into a single, standardized file containing all migration flows for the period 1990-2010. By publishing both the R code to collate, wrangle, and flatten the IRS migration data as well as the migration data itself, my hope is to save time for other migration scholars and facilitate their use. Scholars who wish to use these data should still familiarize themselves with the strengths and weaknesses, idosyncracies, and design of these data (see [@gross2005internal; @engels1981measuring;@franklin2006pandora;@pierce2015soi].

@@

@@

## Setup 
The first part of the code simply sets up the R workspace to facilitate the data wrangling. The approriate R packages are downloaded and installed if the user does not already have these packages installed. The paralell computing environment is also set up as `DetectCores() - 1` to ensure one user core is left for other tasks that computer might need. A single reference tab seperatef (tsv) file is required in this section. `ref_state.tsv` contains FIPS code information for US states. I simply add an additional FIPS code for 'unknown' and assign it FIPS state 99.

## Data Download
The IRS data that can be downloaded is in two primary formats. 1990-2003 and 2004-onwards. This section of code will download and unzip the migration data into a folder standardized format into a subdirectory, `MigData/`. There are eight files that the IRS includes in their zip archives that contain no data. After being downloaded and unzipped, these files are then deleted. If they are not deleted, they will cause the subsequent `for loops` to fail in the next section. These files do not contain any state specific migration information.

## Data Processing
The third and final section contains several `foreach` parallel processing loops to wrangle the seven legacy formats into a common data format. These files are then row-bound using `rbindlist` and transformed into 'short' data frame. **Table 2** demonstrates the general file layout.

# Usage Notes
Any origin-destination pair with fewer than 10 tax filers over the entire period is excluded from the final datafile. This is done to reduce the size of the final datafile.

US Counties are fairly stable geographic units but some changes in counties do exist. To try and keep as close to the original data fidelity as possible, I did not recode any geographic changes. For instance, Broomfield County CO (FIPS 07014) was created out of parts of Adams, Boulder, Jefferson, and Weld counties and thus contains data after 2002. Users should be aware of any changes in geography that could substantially alter any analyses^[More detailed information about county boundary changes can be found at the following locations https://www.census.gov/geo/reference/county-changes.html
http://www.nber.org/asg/ASG_release/County_City/FIPS/FIPS_Changes.pdf
https://www.cdc.gov/nchs/data/nvss/bridged_race/County_Geography_Changes.pdf
https://www.ddorn.net/data/FIPS_County_Code_Changes.pdf].

>>>>

To encourage and facilitate the use of this tremendous migration resource, I flattened the IRS migration data into a single, standardized file containing all migration flows for the period 1990-2010. By publishing both the R code to collate, wrangle, and flatten the IRS migration data as well as the migration data itself, my hope is to save time for other migration scholars and facilitate their use. Scholars who wish to use these data should still familiarize themselves with the strengths and weaknesses, idosyncracies, and design of these data (see [@gross2005internal; @engels1981measuring;@franklin2006pandora;@pierce2015soi] and with the procedures outlined in this document and in the corresponding R code.

@@

Besides the aggregation to FIPS 99999, I performed no other post-processing of the IRS data.

@@

## Setup 
The first part of the code simply sets up the R workspace to facilitate the data wrangling. The approriate R packages are downloaded and installed if the user does not already have these packages installed. The paralell computing environment is also set up as `DetectCores() - 1` to ensure one user core is left for other tasks that computer might need. A single reference tab seperated (tsv) file is required in this section. `ref_state.tsv` contains FIPS code information for US states. I simply add an additional FIPS code for 'unknown' and assign it FIPS state 99.

## Data Download
The IRS data that can be downloaded is in two primary formats. 1990-2003 and 2004-onwards. This section of code will download and unzip the migration data into a folder standardized format into subdirectory, `MigData/`. There are eight files that the IRS includes in their zip archives that contain no data. After being downloaded and unzipped, these files are then deleted. If they are not deleted, they will cause the subsequent `for loops` to fail in the next section. These files do not contain any state specific migration information.

## Data Processing
The third and final section contains several `foreach` parallel processing loops to wrangle the seven legacy formats into a common data format. These files are then row-bound using `rbindlist` and transformed into a 'short' data frame. **Table 2** demonstrates the general file layout.

# Usage Notes
Any origin-destination pair with fewer than 10 tax filers over the entire period is excluded from the final datafile since no data would be recorded in the IRS datafile.

The count data come from the `exemptions` field of the IRS migration data. The original IRS migration data contains two consistent fields across all years of data: a `returns` field and an `exemptions` field. Returns are the actual number of tax returns filed while exemptions are a proxy for the members of the household.

US Counties are fairly stable geographic units but some changes in counties do exist. To try and keep as close to the original data fidelity as possible, I did not recode any geographic changes. For instance, Broomfield County CO (FIPS 07014) was created out of parts of Adams, Boulder, Jefferson, and Weld counties and thus contains data only after 2002. Users should be aware of any changes in geography that could substantially alter any analyses^[More detailed information about county boundary changes can be found at the following locations https://www.census.gov/geo/reference/county-changes.html
http://www.nber.org/asg/ASG_release/County_City/FIPS/FIPS_Changes.pdf
https://www.cdc.gov/nchs/data/nvss/bridged_race/County_Geography_Changes.pdf
https://www.ddorn.net/data/FIPS_County_Code_Changes.pdf].

Users should be aware that I prepared these data to stay as close to the original data fidelity as possible, documented their creation, and with open-source computer code. These data should be used onlly with full awareness of the inherent limitations of the IRS migration data and with the knowledge of the procedures outlined in this document and in the corresponding R code.
